---
title: "ETL : Make tidy data set with "run_analysis.R""
output: github_document
---

# Purpose overview

This readme file have the purpose to describe the methodology and the code steps of **run_analysis.R** file which return in output a tidy data set formed in the text file **avgMeasures.txt**.
To get more infos on the variables output of **avgMeasures.txt**, see **variables_codebook.txt**.
This work is made for the peer-graded assignment course project of the 4th week lesson from the *Getting and Cleaning Data* course by John Hopkins University on coursera plateform.
The **run_analysis.R** code purpose is to extract and transform data from *Human Activity Recognition Using Smartphones Data Set Version 1.0* research [link](https://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones), source :

Jorge L. Reyes-Ortiz, Davide Anguita, Alessandro Ghio, Luca Oneto.
Smartlab - Non Linear Complex Systems Laboratory
DITEN - Università degli Studi di Genova.
Via Opera Pia 11A, I-16145, Genoa, Italy.
activityrecognition@smartlab.ws
www.smartlab.ws

# Summary informations before running the code

## Data

Data folder was download manually as a zip file from the source. Then, it was unzipped manually and directly move in the working directory without any change.

## Running session

```{r}
version
```


```{r}
# Working directory :
getwd()

# Files in working directory :
list.files(getwd())

# Files in the unziped data source :
list.files("./UCI HAR Dataset")
```
# Code steps

## Libraries loaded

The following packages are loaded :

```{r}
library(data.table)  ## Mostly used to read txt with fread() faster than read.table(), but could be replaced easily
library(dplyr)       ## Mostly for merging tables, could be replace with a few more code
library(tidyr)       ## To make tidy a messy data set
library(readr)       ## To parse the numeric from a string, could be easily replace
```

## Reading data

All the data is read from **UCI HAR Dataset** folder and store :

```{r}
datafolder <- "UCI HAR Dataset/"

trainfeature <- fread(file = paste(datafolder, "train/X_train.txt", sep = ""))
trainlabel <- fread(file = paste(datafolder, "train/y_train.txt", sep = ""))
trainsubject <- fread(file = paste(datafolder, "train/subject_train.txt", sep = ""))

testfeature <- fread(file = paste(datafolder, "test/X_test.txt", sep = ""))
testlabel <- fread(file = paste(datafolder, "test/y_test.txt", sep = ""))
testsubject <- fread(file = paste(datafolder, "test/subject_test.txt", sep = ""))

featurenames <- fread(file = paste(datafolder, "features.txt", sep = ""))
labelnames <- fread(file = paste(datafolder, "activity_labels.txt", sep = ""))
```

## Combine train and test datasets (measurements, activity labels and subject ids)

```{r}
combfeature <- bind_rows(trainfeature, testfeature)
comblabels <- bind_rows(trainlabel, testlabel)
combsubject <- bind_rows(trainsubject, testsubject)
```

Quick overview on the merged data sets :

```{r}
dim(combfeature)
dim(comblabels)
dim(combsubject)

options(max.print= 5)
combfeature
```

